Understanding Traffic Patterns and Median Speeds: Insights from Toronto’s Speed Display Program

### 2024-12-01 paper/paper.qmd
added a frequentist model

Speed-related traffic accidents are a major concern in urban areas, prompting cities to explore interventions such as speed display signs to encourage safer driving behavior. The Mobile Watch Your Speed Program (WYSP) in Toronto rotates radar-equipped signs across various locations to remind drivers of their speeds and promote compliance with posted limits. Despite the program’s widespread use, the impact of these installations on traffic speed patterns has not been thoroughly analyzed. This paper investigates the relationship between traffic direction, high-speed vehicle prevalence, and median driving speeds at WYSP locations. By leveraging a Bayesian linear regression model, we aim to uncover patterns that can inform traffic management policies and improve road safety.

@@

@@

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

>>>>

Speeding behavior has long been a focus of traffic safety research, with studies highlighting the psychological and situational factors influencing drivers’ speed choices [@fildes1991speed]. Drivers’ attitudes toward risk, control, and perceived enforcement often play a critical role in determining whether they comply with speed limits or exceed them [@fildes1991speed]. Understanding these attitudes is essential for designing effective interventions that target both behavioral and environmental contributors to speeding [@fildes1991speed]. By situating these insights within the context of Toronto's WYSP, this study explores how structural factors, such as traffic direction, and behavioral tendencies, such as high-speed driving, shape observed speed patterns [@fildes1991speed]. Speed-related traffic accidents are a major concern in urban areas, prompting cities to explore interventions such as speed display signs to encourage safer driving behavior. The Mobile Watch Your Speed Program (WYSP) in Toronto rotates radar-equipped signs across various locations to remind drivers of their speeds and promote compliance with posted limits. Despite the program’s widespread use, the impact of these installations on traffic speed patterns has not been thoroughly analyzed. This paper investigates the relationship between traffic direction, high-speed vehicle prevalence, and median driving speeds at WYSP locations. By leveraging a Bayesian linear regression model, we aim to uncover patterns that can inform traffic management policies and improve road safety.

@@

The chosen predictors, traffic direction and the count of vehicles exceeding 100 km/h, are included based on their theoretical and practical relevance to understanding speed behavior. Traffic direction serves as a proxy for structural and environmental differences between roadways, such as varying traffic density, road design, or signalization, which can influence driving patterns. By using dummy variables for Northbound, Southbound, and Westbound directions with Eastbound as the reference, the model captures these directional variations in median speed, allowing for comparisons across different contexts. The count of vehicles exceeding 100 km/h is included as a continuous predictor to address behavioral influences on median speed. Locations with a higher frequency of speeding vehicles are hypothesized to exhibit elevated central speed tendencies due to social and psychological factors like peer influence or reduced perceived enforcement. Together, these predictors provide a comprehensive framework to disentangle structural and behavioral contributors to speed dynamics, ensuring that the analysis reflects both the physical and social dimensions of driving behavior.

@@

modelsummary::modelsummary(
  list(
    "Bayesian model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```
